---
title: "moth"
author: "Lisa Amalia"
date: "2023-10-26"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#set working directory
```{r}
setwd("G:/My Drive/HUNGARY 2022/3rd SEMESTER/PROJECT EVALUATION AND REVIEW")
```


#A. ESTIMATING POPULATION SIZE AND SURVIVAL RATE

#Load library
```{r}
library(marked)
```

# Load data
```{r}
moth<-read.csv("moth.eps1.csv")
moth <- read.csv("moth.eps1.csv", colClasses = c(ch = "character"))
head(moth)
```

#Process data
```{r}
moth$sex <- as.factor(moth$sex)
moth.js.proc <- process.data(moth, 
                               model = "JS", 
                               groups = "sex")
```
#Make design data from processed data
```{r}
moth.js.ddl <- make.design.data(moth.js.proc)
```

#Fit the data
```{r}
fit.js.moth.models <- function(){
  Phi.dot <- list(formula=~1)
  Phi.time <- list(formula=~time)
  p.dot <- list(formula=~1)
  pent.time <- list(formula=~time)
  pent.sex <- list(formula=~sex)
  pent.dot <- list(formula=~1)
  N.sex <- list(formula=~sex)
  N.dot <- list(formula=~1)
  cml <- create.model.list(c("Phi","p", "pent", "N"))
  results <- crm.wrapper(cml, data = moth.js.proc, ddl = moth.js.ddl,
                         external = FALSE, accumulate = FALSE, hessian = TRUE)
  
  return(results)
}
```

#Run the function
```{r}
moth.js.models <- fit.js.dipper.models()
moth.js.models
```
#Look at the estimated parameters using the top model from the previous code
```{r}
moth.js.models <- fit.js.moth.models()
moth.js.models
moth.js.models[[1]]
```
#The estimates above are not on probability scale (or in individuals for N)

#Look at the prediction for (R E A L) values using top model
```{r}
moth.js.predicted <- predict(moth.js.models[[1]]) # [[1]] just calls the model row according to the model table.
moth.js.predicted #run better from the console
```
#The output shows that:
#survival between capture events is 0.75, 
#detection probability is 0.67, 
#pent is 0.22 each capture event, and 
#the number of unmarked individuals in the superpopulation is about 11 
#(so the super population is ~7 + 294 marked individuals =~ 301). 

#counting the estimated population size
#Abundance (N) is derived from the estimated parameters
#We will estimate population size at each time by making a dataframe of estimates and calculating N
#We will use the predicted estimates from the top-performing model (in this case: "moth.js.predicted")

# NOTE: the below method will have to be adjusted based on your final model and the number of capture events
```{r}
N.derived <- data.frame(occ = c(1:4), #4 events
                        Phi = c(rep(moth.js.predicted$Phi$estimate, 3), NA),   # 3 survival estimates all the same
                        Nsuper = rep(moth.js.predicted$N$estimate + nrow(moth), 4), # Nsuper estimate + number of marked animals
                        pent = c(1-sum(moth.js.predicted$pent$estimate), moth.js.predicted$pent$estimate)) # Sum of all pent must be 1
```

# Set-up empty vector for calculating N
```{r}
N.derived$N <- NA
```

# The inital population size (N[1]) = Nsuper * (1 - sum(all other pent estimates))
# This is because of the link function for estimating pent.
# The sum of all pent parameters MUST equal 1 (therefore, one less must be estimated)
```{r}
N.derived$N[1] <- (N.derived$Nsuper[1] * N.derived$pent[1])
```

# Subsequent population sizes are estimated by calculating surviving individuals (N[t-1] * Phi[t]), and
# Adding new births (Nsuper * pent[t])
```{r}
for(i in 2:nrow(N.derived)){
  N.derived$N[i] <- (N.derived$N[i-1]*N.derived$Phi[i-1]) + (N.derived$Nsuper[i] * N.derived$pent[i])
}
```

# Look at what we did
```{r}
N.derived
```
#B.CHECKING THE CONFIDENCE LEVEL

#Load the library
```{r}
detach("package:marked", unload=TRUE) # many of the function names are the same. unload `marked`
library(RMark)
```
#Process Data. Use the same data but will just create the same top model (not all the other subsets)
```{r}
moth.rmark.processed <- RMark::process.data(dipper,
                                              model = "POPAN")
```

# Formulae for model
```{r}
Phi.dot <- list(formula=~1)
p.dot <- list(formula=~1)
pent.dot <- list(formula=~1)
N.dot <- list(formula=~1)
```



#Load additional library, make sure mark.exe installed on your computer. Fit the data. 
```{r}
library(RMark)
library(TMB)
moth.rmark <- mark(moth, model = "POPAN", 
                     model.parameters = list(Phi = Phi.dot, p= p.dot, 
                                                     pent = pent.dot, N = N.dot),
                     realvcv = TRUE)
```
# The popan.derived function of RMark estimates N 
# (plus estimates SE and 95% CI using the delta method)
```{r}
moth.derived.rmark <- popan.derived(moth.rmark.processed,
                                      moth.rmark)$N
moth.derived.rmark
```

#Plot the result
```{r}
library(ggplot2)

#  first data frame
data1 <- data.frame(
  Occasion = 1:4,
  N = c(118.7933, 163.6329, 199.1518, 227.2875),
  SE = c(23.46395, 18.15735, 20.70510, 27.56353)
)

#  second data frame
data2 <- data.frame(
  occ = 1:4,
  N = c(46.87479, 66.03763, 80.56617, 91.58117)
)

# Create the plot with points, lines, error bars, and customize the theme
plot <- ggplot(data1, aes(x = Occasion, y = N)) +
  geom_point(aes(color = "RMark")) +  # Add points for the first data frame
  geom_line(aes(color = "RMark")) +   # Add lines for the first data frame
  geom_errorbar(aes(ymin = N - SE, ymax = N + SE, color = "RMark"), width = 0.2) +  # Add error bars for the first data frame
  geom_line(data = data2, aes(x = occ, y = N, color = "Hand")) +  # Add a line for the second data frame
  geom_point(data = data2, aes(x = occ, y = N, color = "Hand")) +
  labs(x = "Capture Event", y = "Estimated Population Size") +  # Label the axes
  ggtitle("Estimated Population Size") +  # Add a title
  scale_color_manual(values = c("RMark" = "red", "Hand" = "blue")) +  # Set legend colors
  theme_grey() +  # Use a minimal theme (you can change this to another theme)
  theme(
    plot.title = element_text(size = 16, face = "bold"),  # Customize title text
    axis.text = element_text(size = 12),  # Customize axis text
    axis.title = element_text(size = 14),  # Customize axis title text
    panel.grid.major = element_blank(),  # Remove major grid lines
    panel.grid.minor = element_blank()  # Remove minor grid lines
  )

# Add a legend with custom title and labels
plot<-plot + labs(color = "Estimated by") + 
  scale_color_manual(values = c("RMark" = "red", "Hand" = "blue"),
                    labels = c("RMark" = "RMark", "Hand" = "marked"))
plot + theme(plot.title = element_text(hjust = 0.5, vjust = 0.5))

```
#I choose to display the result with 95% confidence level
```{r}

```


#B.ANALYSING THE HOME-RANGE

# Create a contingency table (matrix) in R
```{r}
contingency_table <- matrix(c(
  105, 55, 2, 1, 0, 2,
  4, 3, 5, 1, 0, 0,
  3, 0, 2, 2, 0, 0
), nrow = 3, byrow = TRUE)  # Set the number of rows and arrange by row

# Assign row and column names
rownames(contingency_table) <- c("Area 1", "Area 2", "Area 3")
colnames(contingency_table) <- c("Male - First", "Female - First", "Male - Second", "Female - Second", "Male - Third", "Female - Third")

# Print the contingency table
contingency_table
```

#Contingency Test
```{r}
# Perform a Chi-Square test
result <- chisq.test(contingency_table)

# Print the Chi-Square test result
print(result)
```

#Analysing The Home Range
```{r}
# Read the data 
hr <- read.csv("moth.hr.csv", header = TRUE)
head(hr)
```
```{r}
# Summary statistics
summary(hr)

# Create cross-tabulations to understand the counts
table(hr$Sex, hr$Capture)
table(hr$Recapture)

```

#Difference between movemon on Male vs Female
```{r}
# Perform a Chi-Square test of independence
chi_square_result <- chisq.test(hr$Sex, hr$Recapture)

# Print the test result
print(chi_square_result)

```
#There is a significat differences between Male and Female Movement. Male more active

#Different Movement on Area 1, Area 2, Area 3
```{r}
# Load the 'dplyr' package for data manipulation
library(dplyr)

# Group by 'Sex,' 'Capture,' and 'Recapture' and count the occurrences
hr <- hr %>%
  group_by(Sex, Capture, Recapture) %>%
  mutate(Visit = n())

# View the resulting dataset
print(hr)

```

```{r}
# Perform one-way ANOVA (Assume this data not normally distributed)
anova_result <- aov(Visit ~ Recapture, data = hr)

# Print the ANOVA result
summary(anova_result)
```
```{r}
# However, you can load it just to make sure it's available
library(stats)

hr <- hr %>%
  group_by(Sex, Capture, Recapture) %>%
  mutate(Visit = n())
```
#There are significant differences of MOvement in each Area where Moths are mostly move in Area 1

```{r}
# Perform pairwise t-tests with Bonferroni correction
pairwise_t_test_result <- pairwise.t.test(hr$Visit, hr$Recapture, p.adjust.method = "holm")

# View the pairwise comparison results
print(pairwise_t_test_result)
```
# Load the 'dplyr' package
library(dplyr)

# Filter the data frame to select only male individuals and count them
male_count <- df %>% filter(sex == "Male") %>% nrow()

# Print the count of male individuals
cat("Number of male individuals:", male_count, "\n")

