---
title: "moth"
author: "Lisa Amalia"
date: "2023-10-26"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##### 1. General Analysis ----------------------------------

#Load library
```{r}
library(marked)
library(ggplot2)
```

# Load data
```{r}
moth<-read.csv("moth.eps1.csv")
moth <- read.csv("moth.eps1.csv", colClasses = c(ch = "character"))
head(moth)
```

#Process data
```{r}
moth$sex <- as.factor(moth$sex)
moth.js.proc <- process.data(moth, 
                               model = "JS", 
                               groups = "sex")
```
#Make design data from processed data
```{r}
moth.js.ddl <- make.design.data(moth.js.proc)
```

#Fit the data
```{r}
fit.js.moth.models <- function(){
  Phi.dot <- list(formula=~1)
  p.dot <- list(formula=~1)
  p.time <- list(formula=~time)
  p.sex <- list(formula=~sex)
  pent.dot <- list(formula=~1)
  pent.time <- list(formula=~time)
  pent.sex <- list(formula=~sex)
  N.dot <- list(formula=~1)
  N.sex <- list(formula=~sex)
  cml <- create.model.list(c("Phi","p", "pent", "N"))
  results <- crm.wrapper(cml, data = moth.js.proc, ddl = moth.js.ddl,
                         external = FALSE, accumulate = FALSE, hessian = TRUE)
  
  return(results)
}
```

#Run the function
```{r}
moth.js.models <- fit.js.moth.models()
moth.js.models
moth.js.models[[1]] 
```
#The estimates above are not on probability scale (or in individuals for N)
# (e.g. Phi, p on logit scale, pent on mlogit scale)
#Look at the prediction for (real) values using top model
```{r}
moth.js.predicted <- predict(moth.js.models[[1]], ddl = moth.js.ddl, se = TRUE) # [[1]] just calls the model row according to the model table.
moth.js.predicted #run better from the console
```
#The output shows that:
#survival between capture events is 0.75, 
#detection probability is 0.67, 
#pent is 0.22 each capture event, and 
#the number of unmarked individuals in the superpopulation is about 11 
#(so the super population is ~11 + 128 marked individuals =~ 139). 

#counting the estimated population size
#Abundance (N) is derived from the estimated parameters
#We will estimate population size at each time by making a dataframe of estimates and calculating N
#We will use the predicted estimates from the top-performing model (in this case: "moth.js.predicted")

# NOTE: the below method will have to be adjusted based on your final model and the number of capture events
```{r}
N.derived <- data.frame(occ = c(1:4), #4 events
                        Phi = c(rep(moth.js.predicted$Phi$estimate, 3), NA),   # 3 survival estimates all the same
                        Nsuper = rep(moth.js.predicted$N$estimate + nrow(moth), 4), # Nsuper estimate + number of marked animals
                        pent = c(1-sum(moth.js.predicted$pent$estimate), moth.js.predicted$pent$estimate)) # Sum of all pent must be 1
```

# Set-up empty vector for calculating N
```{r}
N.derived$N <- NA
N.derived$N.lcl <- NA
N.derived$N.ucl <- NA
```

# The inital population size (N[1]) = Nsuper * (1 - sum(all other pent estimates))
# This is because of the link function for estimating pent.
# The sum of all pent parameters MUST equal 1 (therefore, one less must be estimated)
```{r}
N.derived$N[1] <- (N.derived$Nsuper[1] * N.derived$pent[1])
# Adding SE / confidence intervals around population size estimates
N.derived$N.ucl[1] <- (moth.js.predicted$N$ucl + nrow(moth))*(1-sum(moth.js.predicted$pent$lcl))
N.derived$N.lcl[1] <- (moth.js.predicted$N$lcl + nrow(moth))*(1-sum(moth.js.predicted$pent$ucl))
```

# Subsequent population sizes are estimated by calculating surviving individuals (N[t-1] * Phi[t]), and
# Adding new births (Nsuper * pent[t])
```{r}
for(i in 2:nrow(N.derived)){
  N.derived$N[i] <- (N.derived$N[i-1]*N.derived$Phi[i-1]) + (N.derived$Nsuper[i] * N.derived$pent[i])
  N.derived$N.ucl[i] <- (N.derived$N.ucl[i-1]*moth.js.predicted$Phi$ucl[1]) + (((moth.js.predicted$N$ucl + nrow(moth))) *moth.js.predicted$pent$ucl[i-1])
  N.derived$N.lcl[i] <- (N.derived$N.lcl[i-1]*moth.js.predicted$Phi$lcl[1]) + (((moth.js.predicted$N$lcl + nrow(moth))) *moth.js.predicted$pent$lcl[i-1])
}
```

# The final result
```{r}
N.derived
```

#Plot the estimated poppulation size
```{r}
ggplot(N.derived, aes(x = occ, y = N)) +
  geom_point(shape = 16, size = 3, col = "black") +
  geom_errorbar(aes(ymin = N.lcl, ymax = N.ucl), width = 0.3) +
  geom_path() +
  labs(x = "Day", y = "Estimated population size(overall)\n(and 95% confidence interval)") +
  theme_grey()
```

##### 2. Analysis for Female ----------------------------------

```{r}
moth_female <- subset(moth, sex == "Female")
mothfemale.js.proc <- process.data(moth_female, model = "JS")
mothfemale.js.ddl <- make.design.data(mothfemale.js.proc)
fit.js.mothfemale.models <- function(){
  Phi.dot <- list(formula=~1)
  p.dot <- list(formula=~1)
  p.time <- list(formula=~time)
  pent.dot <- list(formula=~1)
  pent.time <- list(formula=~time)
  N.dot <- list(formula=~1)
  cml <- create.model.list(c("Phi","p", "pent", "N"))
  results <- crm.wrapper(cml, data = mothfemale.js.proc, ddl = mothfemale.js.ddl,
                         external = FALSE, accumulate = FALSE, hessian = TRUE)
  return(results)
}
mothfemale.js.models <- fit.js.mothfemale.models()
mothfemale.js.predicted <- predict(mothfemale.js.models[[1]], ddl = mothfemale.js.ddl, se = TRUE)
```

```{r}
N.derived <- data.frame(occ = c(1:4), #4 events
                        Phi = c(rep(mothfemale.js.predicted$Phi$estimate, 3), NA),   # 3 survival estimates all the same
                        Nsuper = rep(mothfemale.js.predicted$N$estimate + nrow(moth), 4), # Nsuper estimate + number of marked animals
                        pent = c(1-sum(mothfemale.js.predicted$pent$estimate), mothfemale.js.predicted$pent$estimate)) # Sum of all pent must be 1
```

# Set-up empty vector for calculating N
```{r}
N.derived$N <- NA
N.derived$N.lcl <- NA
N.derived$N.ucl <- NA
```

# The inital population size (N[1]) = Nsuper * (1 - sum(all other pent estimates))
# This is because of the link function for estimating pent.
# The sum of all pent parameters MUST equal 1 (therefore, one less must be estimated)
```{r}
N.derived$N[1] <- (N.derived$Nsuper[1] * N.derived$pent[1])
# Adding SE / confidence intervals around population size estimates
N.derived$N.ucl[1] <- (mothfemale.js.predicted$N$ucl + nrow(moth))*(1-sum(mothfemale.js.predicted$pent$lcl))
N.derived$N.lcl[1] <- (mothfemale.js.predicted$N$lcl + nrow(moth))*(1-sum(mothfemale.js.predicted$pent$ucl))
```

# Subsequent population sizes are estimated by calculating surviving individuals (N[t-1] * Phi[t]), and
# Adding new births (Nsuper * pent[t])
```{r}
for(i in 2:nrow(N.derived)){
  N.derived$N[i] <- (N.derived$N[i-1]*N.derived$Phi[i-1]) + (N.derived$Nsuper[i] * N.derived$pent[i])
  N.derived$N.ucl[i] <- (N.derived$N.ucl[i-1]*mothfemale.js.predicted$Phi$ucl[1]) + (((mothfemale.js.predicted$N$ucl + nrow(moth))) *mothfemale.js.predicted$pent$ucl[i-1])
  N.derived$N.lcl[i] <- (N.derived$N.lcl[i-1]*mothfemale.js.predicted$Phi$lcl[1]) + (((mothfemale.js.predicted$N$lcl + nrow(moth))) *mothfemale.js.predicted$pent$lcl[i-1])
}
```

# The final result
```{r}
N.derived
```
#Plot the estimated poppulation size
```{r}
ggplot(N.derived, aes(x = occ, y = N)) +
  geom_point(shape = 16, size = 3, col = "black") +
  geom_errorbar(aes(ymin = N.lcl, ymax = N.ucl), width = 0.3) +
  geom_path() +
  labs(x = "Day", y = "Estimated population size for Female\n(and 95% confidence interval)") +
  theme_grey()
```

##### 3. Analysis for Male ----------------------------------
```{r}
moth_male <- subset(moth, sex == "Male")
mothmale.js.proc <- process.data(moth_male, model = "JS")
mothmale.js.ddl <- make.design.data(mothmale.js.proc)
fit.js.mothmale.models <- function(){
  Phi.dot <- list(formula=~1)
  p.dot <- list(formula=~1)
  p.time <- list(formula=~time)
  pent.dot <- list(formula=~1)
  pent.time <- list(formula=~time)
  N.dot <- list(formula=~1)
  cml <- create.model.list(c("Phi","p", "pent", "N"))
  results <- crm.wrapper(cml, data = mothmale.js.proc, ddl = mothmale.js.ddl,
                         external = FALSE, accumulate = FALSE, hessian = TRUE)
  return(results)
}
mothmale.js.models <- fit.js.mothmale.models()
mothmale.js.predicted <- predict(mothmale.js.models[[1]], ddl = mothmale.js.ddl, se = TRUE)
```

```{r}
N.derived <- data.frame(occ = c(1:4), #4 events
                        Phi = c(rep(mothmale.js.predicted$Phi$estimate, 3), NA),   # 3 survival estimates all the same
                        Nsuper = rep(mothmale.js.predicted$N$estimate + nrow(moth), 4), # Nsuper estimate + number of marked animals
                        pent = c(1-sum(mothmale.js.predicted$pent$estimate), mothmale.js.predicted$pent$estimate)) # Sum of all pent must be 1
```

# Set-up empty vector for calculating N
```{r}
N.derived$N <- NA
N.derived$N.lcl <- NA
N.derived$N.ucl <- NA
```

# The inital population size (N[1]) = Nsuper * (1 - sum(all other pent estimates))
# This is because of the link function for estimating pent.
# The sum of all pent parameters MUST equal 1 (therefore, one less must be estimated)
```{r}
N.derived$N[1] <- (N.derived$Nsuper[1] * N.derived$pent[1])
# Adding SE / confidence intervals around population size estimates
N.derived$N.ucl[1] <- (mothmale.js.predicted$N$ucl + nrow(moth))*(1-sum(mothmale.js.predicted$pent$lcl))
N.derived$N.lcl[1] <- (mothmale.js.predicted$N$lcl + nrow(moth))*(1-sum(mothmale.js.predicted$pent$ucl))
```

# Subsequent population sizes are estimated by calculating surviving individuals (N[t-1] * Phi[t]), and
# Adding new births (Nsuper * pent[t])
```{r}
for(i in 2:nrow(N.derived)){
  N.derived$N[i] <- (N.derived$N[i-1]*N.derived$Phi[i-1]) + (N.derived$Nsuper[i] * N.derived$pent[i])
  N.derived$N.ucl[i] <- (N.derived$N.ucl[i-1]*mothmale.js.predicted$Phi$ucl[1]) + (((mothmale.js.predicted$N$ucl + nrow(moth))) *mothmale.js.predicted$pent$ucl[i-1])
  N.derived$N.lcl[i] <- (N.derived$N.lcl[i-1]*mothmale.js.predicted$Phi$lcl[1]) + (((mothmale.js.predicted$N$lcl + nrow(moth))) *mothmale.js.predicted$pent$lcl[i-1])
}
```

# The final result
```{r}
N.derived
```
#Plot the estimated poppulation size
```{r}
ggplot(N.derived, aes(x = occ, y = N)) +
  geom_point(shape = 16, size = 3, col = "black") +
  geom_errorbar(aes(ymin = N.lcl, ymax = N.ucl), width = 0.3) +
  geom_path() +
  labs(x = "Day", y = "Estimated population size for Male\n(and 95% confidence interval)") +
  theme_grey()
```

##### 4. Analysis for Male ----------------------------------

# Create a contingency table (matrix) in R
```{r}
contingency_table <- matrix(c(
  105, 55, 2, 1, 0, 2,
  4, 3, 5, 1, 0, 0,
  3, 0, 2, 2, 0, 0
), nrow = 3, byrow = TRUE)  # Set the number of rows and arrange by row

# Assign row and column names
rownames(contingency_table) <- c("Area 1", "Area 2", "Area 3")
colnames(contingency_table) <- c("Male - First", "Female - First", "Male - Second", "Female - Second", "Male - Third", "Female - Third")

# Print the contingency table
contingency_table
```

#Contingency Test
```{r}
# Perform a Chi-Square test
result <- chisq.test(contingency_table)

# Print the Chi-Square test result
print(result)
```

#Analysing The Home Range
```{r}
# Read the data 
hr <- read.csv("moth.hr290ct.csv", header = TRUE)
head(hr)
```

citation()
